---
title: "Week5_COVID19_Analysis"
output: pdf_document
date: "2022-08-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
```

# Summary

This is an analysis on COVID-19 data from Johns Hopkins. The data is retrieved from https://github.com/CSSEGISandData/COVID-19.
In this study, I examined the differences between Japan and the UK in terms of the new cases and deaths.
Until recently, Japan is one of the countries that apply strong policies to restrain COVID-19 cases. 
On the other hand, the UK adopts light restrictions to save the economy.
The comparison implies that COVID-19 cannot be zero, and the number of cases/deaths surges whenever the restriction is lifted.

# Import Data

Here is how to read the data. All files are CSV format. So, I use read_csv to read the data. 
```{r importdata}
url_in <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/"
file_names <- c("time_series_covid19_confirmed_global.csv",
                 "time_series_covid19_deaths_global.csv"
                 )
urls <- str_c(url_in, file_names)
uid_lookup_url <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/UID_ISO_FIPS_LookUp_Table.csv"

global_cases <- read_csv(urls[1], show_col_types = FALSE)
global_deaths <- read_csv(urls[2], show_col_types = FALSE)
uid <- read_csv(uid_lookup_url) 
```

# Tidy the data

For this analysis, I do not use latitude and longitude because I will not use them. Then, change the date rows to columns. 
```{r tidy}
global_cases <- global_cases %>% select(-c(`Province/State`, Lat, Long))
global_cases <- global_cases %>% pivot_longer(cols = -c(`Country/Region`), names_to = "date", values_to = "cases")
```
The date column type is not Date object but Char. Let's convert is to date type. 
```{r summary}
summary(global_cases)
global_cases <- global_cases %>% mutate(date = mdy(date))
summary(global_cases)
```

Now, it is clear that the data contains the data from 2020-01-22 to 2022-08-12.

I'll do the same to another data.
```{r tidy_death}
global_deaths <- global_deaths %>% select(-c(`Province/State`, Lat, Long))
global_deaths <- global_deaths %>% pivot_longer(cols = -c(`Country/Region`), names_to = "date", values_to = "deaths")
global_deaths <- global_deaths %>% mutate(date = mdy(date))
```

I will combine those two data as a data frame called global. For convenience, I'll rename the colums as well.
```{r global}
global <- global_cases %>% full_join(global_deaths)
global <- global %>% rename(Country_Region = `Country/Region`)
```

For the later analysis, filter the date by cases more than 0.
```{r summary_global}
global <- global %>% filter(cases>0)
```
Add the population to the data frame. Now, I have the data with number of cases, deaths, and population in each country per day from 2020-01-22 to 2022-08-12.
```{r population}
uid <- uid %>% select(-c(Lat, Long_, Combined_Key, code3, iso2, iso3, Admin2))
global <- global %>% left_join(uid, by = c("Country_Region")) 
global <- global %>% select(-c(UID, FIPS)) 
global <- global %>% select(Country_Region, date, cases, deaths, Population)
summary(global)
```
As the last step, add the number of new cases/deaths.
```{r per_population}
global <- global %>% group_by(Country_Region, date)
global <- global %>% summarise(cases = sum(cases), deaths = sum(deaths), Population = sum(Population))
global <- global %>% mutate(new_cases = cases - lag(cases), new_deaths = deaths - lag(deaths))
```

# Analysis

Here is the number of new cases and new deaths in Japan. Japan kept the number of deaths low. 
In Oct 2021, Kishida became the prime minister. He ceased a number of restrictions to rescue the economy.
That results in a rapid increase of the new cases. 
Although the number of death is still relatively low, it will increase since the peak of deaths will come after the peak of cases.

```{r japan}
japan <- global[global$Country_Region=="Japan",] %>% filter(cases>0) %>% filter(new_cases>0) %>% filter(deaths>0) %>% filter(new_deaths > 0)
japan %>% ggplot(aes(x = date, y = cases)) + geom_line(aes(color = "cases")) + geom_line(aes(y = deaths, color = "deaths")) + scale_y_log10()
japan %>% ggplot(aes(x = date, y = new_cases)) + geom_line(aes(color = "new_cases")) + geom_line(aes(y = new_deaths, color = "new_deaths")) + scale_y_log10()
```
The UK went through a completely different path. UK government tried to remove the restriction earlier to save the economy.
As a result, the number of cases and deaths are much higher than in Japan, even though the population in the UK is half of Japan's population.
The number of cases and deaths has kept being high. It seems the UK accepted the risk of losing their health.

```{r uk}
uk <- global[global$Country_Region=="United Kingdom",] %>% filter(cases>0) %>% filter(new_cases>0) %>% filter(deaths>0) %>% filter(new_deaths > 0) 
uk %>% ggplot(aes(x = date, y = cases)) + geom_line(aes(color = "cases")) + geom_point(aes(color = "cases")) + geom_line(aes(y = deaths, color = "deaths")) + geom_point(aes(y = deaths, color = "deaths")) + scale_y_log10()
uk %>% ggplot(aes(x = date, y = new_cases)) + geom_line(aes(color = "new_cases")) + geom_line(aes(y = new_deaths, color = "new_deaths")) + scale_y_log10()
```

# Conclusion

In Japan, the analysis shows that the lift of the restriction causes an increase in cases.
Japan has adopted the "hammer and dance" strategy against COVID-19.
That resulted in multiple waves of new cases.
It is clear that there is a correlation between the government restriction policy and the number of cases.

I have been biased that the UK (and other countries) applied strong movement restrictions (lockdown) as Japan had at least at the beginning of the pandemic.
But according to the historical data of cases and deaths, the restriction would not be effective in the UK as I thought it was.

Another bias I had is related to the number of waves. 
I assumed the UK also changed policies several times as Japan did. But it is not clear from the graph.
The reason would be either change of the restriction didn't affect the people's behavior, or the government policy has been relatively consistent in the last two years.

I speculate that people in the UK have a stronger belief in the freedom of mobility.
Under that situation, the strong restriction only hurts the economy and doesn't work effectively to suppress COVID-19 cases.
It is important to understand how people react to the policy so that we can have a healthy and safe life.